### 2021-05-27 ssp_example.tex

As an illustrative example, we consider a multistratum experimental framework with three strata in total, and $2$, $1$ and $2$ factors applied in each of them respectively. $48$ experimental unites are set up in $12$ whole-plots, each of them containing $2$ sub-plots with $2$ runs per sub-plot. The fitted full quadratic polynomial model in this case contains $20$ parameters: $5$ in the first stratum, $4$ in the second and $11$ in the third. As for the potential terms, there are $30$ of them: just $2$ and $5$ are  in the whole- and sub-plots, the other $23$ are the third-order terms formed from the factors in the sub-sub-plots and their interactions with the terms from the higher strata.

\begin{table}[h]
\centering
\caption{Summary of MSE(DP)-optimal split-split-plot designs}
\label{tab::mseD_ex2}
%\scalebox{0.75}{
\resizebox{\textwidth}{!}{%
\begin{tabular}{rrrrccccccrrrrr}
  & \multicolumn{3}{l}{\textbf{Criteria, $\bm{\tau^2=1/q}$}}    & \multicolumn{3}{l}{\textbf{DF (s=1)}}   & \multicolumn{3}{l}{\textbf{DF (s=2)}}    & \multicolumn{2}{l}{\textbf{DF(s=3)}} & \multicolumn{3}{l}{\textbf{Efficiencies,\%}}     \\
  & \textbf{DP} & \textbf{LoF(DP)} & \textbf{MSE(D)} & \textbf{PE} & \textbf{Inter-WP} & \textbf{LoF} & \textbf{PE} & \textbf{Inter-SP} & \textbf{LoF}& \textbf{PE}      & \textbf{LoF}      & \textbf{DP} & \textbf{LoF(DP)} & \textbf{MSE(D)} \\
1 & 1    & 0    & 0    & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{5} & 1 & 2 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{100.00} & 91.44  & 85.32  \\
2 & 0    & 1    & 0    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{3} & 5 & 0 & \multicolumn{1}{|r}{13} & 0  & \multicolumn{1}{|r}{5.99}   & 100.00 & 5.06   \\
3 & 0    & 0    & 1    & \multicolumn{1}{|r}{0} & 3 & 3 & \multicolumn{1}{|r}{0} & 3 & 5 & \multicolumn{1}{|r}{0}  & 13 & \multicolumn{1}{|r}{0.00}  & 0.00  & 100.00 \\
4 &1/3  & 1/3  & 1/3   & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{3} & 3 & 2 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{92.85}  & 87.17  & 83.81  \\
5 & 0.5  & 0.5  & 0    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{6} & 1 & 1 & \multicolumn{1}{|r}{11} & 2  & \multicolumn{1}{|r}{84.77}  & 96.02  & 70.81  \\
6 & 0.5  & 0    & 0.5  & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{4} & 1 & 3 & \multicolumn{1}{|r}{8}  & 5  & \multicolumn{1}{|r}{93.22}  & 83.11  & 90.42  \\
7 & 0    & 0.5  & 0.5  & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{2} & 4 & 2 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{85.98}  & 90.59  & 79.33  \\
8 & 0.5  & 0.25 & 0.25 & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{4} & 2 & 2 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{90.88}  & 91.03  & 83.39  \\
  &      &      &      &   &   &   &   &   &   &    &    &        &        &        \\  
  & \multicolumn{3}{l}{\textbf{Criteria, $\bm{\tau^2=\sqrt{1/q}}$}}    & \multicolumn{3}{l}{\textbf{DF (s=1)}}   & \multicolumn{3}{l}{\textbf{DF (s=2)}}    & \multicolumn{2}{l}{\textbf{DF(s=3)}} & \multicolumn{3}{l}{\textbf{Efficiencies,\%}}     \\
  & \textbf{DP} & \textbf{LoF(DP)} & \textbf{MSE(D)} & \textbf{PE} & \textbf{Inter-WP} & \textbf{LoF} & \textbf{PE} & \textbf{Inter-SP} & \textbf{LoF}& \textbf{PE}      & \textbf{LoF}      & \textbf{DP} & \textbf{LoF(DP)} & \textbf{MSE(D)} \\
1 & 1    & 0    & 0    & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{5} & 1 & 2 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{100.00} & 92.11  & 84.07  \\
2 & 0    & 1    & 0    & \multicolumn{1}{|r}{1} & 2 & 3 & \multicolumn{1}{|r}{2} & 5 & 1 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{30.24}  & 100.00 & 28.05  \\
3 & 0    & 0    & 1    & \multicolumn{1}{|r}{0} & 3 & 3 & \multicolumn{1}{|r}{0} & 3 & 5 & \multicolumn{1}{|r}{0}  & 13 & \multicolumn{1}{|r}{0.00}   & 0.00   & 100.00 \\
4 &1/3  & 1/3  & 1/3   & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{5} & 1 & 2 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{86.39}  & 96.27  & 78.05  \\
5 & 0.5  & 0.5  & 0    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{3} & 3 & 2 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{83.93}  & 100.76 & 76.45  \\
6 & 0.5  & 0    & 0.5  & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{5} & 0 & 3 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{96.90}  & 90.06  & 86.09  \\
7 & 0    & 0.5  & 0.5  & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{2} & 3 & 3 & \multicolumn{1}{|r}{8}  & 5  & \multicolumn{1}{|r}{77.78}  & 97.70  & 78.21  \\
8 & 0.5  & 0.25 & 0.25 & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{2} & 4 & 2 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{88.17}  & 95.22  & 79.23  \\
  &      &      &      &   &   &   &   &   &   &    &    &        &        &        \\  
  & \multicolumn{3}{l}{\textbf{Criteria, $\bm{\tau^2=1}$}}    & \multicolumn{3}{l}{\textbf{DF (s=1)}}   & \multicolumn{3}{l}{\textbf{DF (s=2)}}    & \multicolumn{2}{l}{\textbf{DF(s=3)}} & \multicolumn{3}{l}{\textbf{Efficiencies,\%}}     \\
  & \textbf{DP} & \textbf{LoF(DP)} & \textbf{MSE(D)} & \textbf{PE} & \textbf{Inter-WP} & \textbf{LoF} & \textbf{PE} & \textbf{Inter-SP} & \textbf{LoF}& \textbf{PE}      & \textbf{LoF}      & \textbf{DP} & \textbf{LoF(DP)} & \textbf{MSE(D)} \\
1 & 1    & 0    & 0    & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{5} & 1 & 2 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{100.00} & 69.23  & 89.76  \\
2 & 0    & 1    & 0    & \multicolumn{1}{|r}{2} & 2 & 2 & \multicolumn{1}{|r}{1} & 4 & 3 & \multicolumn{1}{|r}{7}  & 6  & \multicolumn{1}{|r}{46.43}  & 100.00 & 51.81  \\
3 & 0    & 0    & 1    & \multicolumn{1}{|r}{1} & 2 & 3 & \multicolumn{1}{|r}{0} & 3 & 5 & \multicolumn{1}{|r}{0}  & 13 & \multicolumn{1}{|r}{0.00}   & 0.00   & 100.00 \\
4 &1/3  & 1/3  & 1/3   & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{5} & 0 & 3 & \multicolumn{1}{|r}{8}  & 5  & \multicolumn{1}{|r}{83.38}  & 84.60  & 88.46  \\
5 & 0.5  & 0.5  & 0    & \multicolumn{1}{|r}{1} & 2 & 3 & \multicolumn{1}{|r}{3} & 2 & 3 & \multicolumn{1}{|r}{8}  & 5  & \multicolumn{1}{|r}{74.06}  & 87.14  & 78.63  \\
6 & 0.5  & 0    & 0.5  & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{5} & 0 & 3 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{98.09}  & 74.68  & 93.25  \\
7 & 0    & 0.5  & 0.5  & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{3} & 2 & 3 & \multicolumn{1}{|r}{6}  & 7  & \multicolumn{1}{|r}{81.25}  & 94.80  & 93.70  \\
8 & 0.5  & 0.25 & 0.25 & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{2} & 3 & 3 & \multicolumn{1}{|r}{8}  & 5  & \multicolumn{1}{|r}{83.00}  & 87.46  & 87.02  
\end{tabular}
}
\end{table}

The summaries of $MSE(DP)$- and $MSE(LP)$-optimal designs' performances and the distributions of the degrees of freedom are given in Tables \ref{tab::mseD_ex2} and \ref{tab::mseL_ex2} respectively. In the first two strata some of the degrees of freedom are allocated to the inter-whole-plot (``Inter-WP'') and inter-sub-plot (``Inter-SP'') information components.

Even though there are not many residual degrees of freedom available at each stratum, the pattern of the imbalance between the pure error and lack-of-fit components can be observed: lower values of the variance scaling parameter $\tau^2$ result in more replicate. In the case of $MSE(LP)$-optimal designs this pattern seems to be coherent with the weight put on the $LP$- and $LoF(LP)$-components.

In general, similar to the examples seen before, searching for compound optimality results in designs performing decently in terms of $DP$-, $LP$- and lack-of-fit components, with the efficiencies dropping when no weight is allocated to the corresponding part of the criterion and/or with the value of $\tau^2$ increasing. Increased variance scaling parameter - meaning larger prior scale of the fitted model disturbance - leads to generally higher $MSE(D)$-efficiency of the optimal designs, that is protecting the quality of the primary model parameters' estimates from the potential deviation.

Most of the compound-optimal designs can provide a satisfying compromise among the three objectives; however, $MSE(D)-/MSE(L)-$ efficiency values are more responsive to the allocation of weights; so once an MSE-based component is of importance, this tendency certainly needs to be taken into account regardless of what criterion and variance scaling parameter is used.

An interesting observation worth noting is the $MSE(DP)$-optimal design \#$5$ (for $\tau^2=\sqrt{1/q}$) that was obtained as $DP$- and $LoF(DP)$-efficient turned out to be slightly more $LoF(DP)$-efficient ($100.76\%$) rather than the one that was searched for as $LoF(DP)$-optimal (\#$2$); besides, it performs much better in terms of the other criterion components, i.e. it is more than $70\%$ efficient w.r.t. each of them. This is not something that would be normally expected, but the original design search procedure is expected to find near-optimum designs, especially when the expectation in the MSE(DP)-criterion is approximated -- and that holds for both unblocked, blocked and multistratum frameworks.

%%%% Trace-based, Example 2

\begin{table}[h]
\centering
\caption{Example 2. Properties of MSE(L)-optimal designs}
\label{tab::mseL_ex2}
%\scalebox{0.75}{
\resizebox{\textwidth}{!}{%
\begin{tabular}{rrrrccccccrrrrr}
  & \multicolumn{3}{l}{\textbf{Criteria, $\bm{\tau^2=1/q}$}}    & \multicolumn{3}{l}{\textbf{DF (s=1)}}   & \multicolumn{3}{l}{\textbf{DF (s=2)}}    & \multicolumn{2}{l}{\textbf{DF(s=3)}} & \multicolumn{3}{l}{\textbf{Efficiencies,\%}}     \\
  & \textbf{LP} & \textbf{LoF(LP)} & \textbf{MSE(L)} & \textbf{PE} & \textbf{Inter-WP} & \textbf{LoF} & \textbf{PE} & \textbf{Inter-SP} & \textbf{LoF}& \textbf{PE}      & \textbf{LoF}      & \textbf{LP} & \textbf{LoF(LP)} & \textbf{MSE(L)} \\
1 & 1    & 0    & 0    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{3} & 3 & 2 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{100.00} & 84.14  & 53.61  \\
2 & 0    & 1    & 0    & \multicolumn{1}{|r}{1} & 2 & 3 & \multicolumn{1}{|r}{3} & 5 & 0 & \multicolumn{1}{|r}{13} & 0  & \multicolumn{1}{|r}{0.00}   & 100.00 & 0.00   \\
3 & 0    & 0    & 1    & \multicolumn{1}{|r}{0} & 3 & 3 & \multicolumn{1}{|r}{0} & 1 & 7 & \multicolumn{1}{|r}{0}  & 13 & \multicolumn{1}{|r}{0.00}   & 0.00   & 100.00 \\
4 & 1/3 & 1/3 & 1/3    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{4} & 2 & 2 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{88.77}  & 84.09  & 59.95  \\
5 & 0.5  & 0.5  & 0    & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{2} & 5 & 1 & \multicolumn{1}{|r}{12} & 1  & \multicolumn{1}{|r}{94.41}  & 95.38  & 48.74  \\
6 & 0.5  & 0    & 0.5  & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{2} & 3 & 3 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{89.77}  & 76.14  & 67.58  \\
7 & 0    & 0.5  & 0.5  & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{3} & 2 & 3 & \multicolumn{1}{|r}{11} & 2  & \multicolumn{1}{|r}{81.78}  & 91.51  & 59.73  \\
8 & 0.5  & 0.25 & 0.25 & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{2} & 4 & 2 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{95.82}  & 83.35  & 57.59  \\
  &      &      &      &   &   &   &   &   &   &    &    &        &        &        \\  
  & \multicolumn{3}{l}{\textbf{Criteria, $\bm{\tau^2=\sqrt{1/q}}$}}    & \multicolumn{3}{l}{\textbf{DF (s=1)}}   & \multicolumn{3}{l}{\textbf{DF (s=2)}}    & \multicolumn{2}{l}{\textbf{DF(s=3)}} & \multicolumn{3}{l}{\textbf{Efficiencies,\%}}     \\
  & \textbf{LP} & \textbf{LoF(LP)} & \textbf{MSE(L)} & \textbf{PE} & \textbf{Inter-WP} & \textbf{LoF} & \textbf{PE} & \textbf{Inter-SP} & \textbf{LoF}& \textbf{PE}      & \textbf{LoF}      & \textbf{LP} & \textbf{LoF(LP)} & \textbf{MSE(L)} \\
1 & 1    & 0    & 0    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{3} & 3 & 2 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{100.00} & 88.95  & 31.38  \\
2 & 0    & 1    & 0    & \multicolumn{1}{|r}{0} & 3 & 3 & \multicolumn{1}{|r}{2} & 5 & 1 & \multicolumn{1}{|r}{13} & 0  & \multicolumn{1}{|r}{0.00}   & 100.00 & 0.00   \\
3 & 0    & 0    & 1    & \multicolumn{1}{|r}{0} & 3 & 3 & \multicolumn{1}{|r}{0} & 0 & 8 & \multicolumn{1}{|r}{0}  & 13 & \multicolumn{1}{|r}{0.00}   & 0.00   & 100.00 \\
4 & 1/3 & 1/3 & 1/3    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{2} & 4 & 2 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{85.15}  & 88.87  & 39.06  \\
5 & 0.5  & 0.5  & 0    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{2} & 5 & 1 & \multicolumn{1}{|r}{11} & 2  & \multicolumn{1}{|r}{91.36}  & 93.49  & 31.90  \\
6 & 0.5  & 0    & 0.5  & \multicolumn{1}{|r}{3} & 0 & 3 & \multicolumn{1}{|r}{3} & 2 & 3 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{93.18}  & 82.34  & 41.75  \\
7 & 0    & 0.5  & 0.5  & \multicolumn{1}{|r}{1} & 2 & 3 & \multicolumn{1}{|r}{1} & 4 & 3 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{57.49}  & 81.98  & 53.67  \\
8 & 0.5  & 0.25 & 0.25 & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{3} & 3 & 2 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{94.10}  & 83.43  & 38.75  \\
  &      &      &      &   &   &   &   &   &   &    &    &        &        &   \\  
  & \multicolumn{3}{l}{\textbf{Criteria, $\bm{\tau^2=1}$}}    & \multicolumn{3}{l}{\textbf{DF (s=1)}}   & \multicolumn{3}{l}{\textbf{DF (s=2)}}    & \multicolumn{2}{l}{\textbf{DF(s=3)}} & \multicolumn{3}{l}{\textbf{Efficiencies,\%}}     \\
  & \textbf{LP} & \textbf{LoF(LP)} & \textbf{MSE(L)} & \textbf{PE} & \textbf{Inter-WP} & \textbf{LoF} & \textbf{PE} & \textbf{Inter-SP} & \textbf{LoF}& \textbf{PE}      & \textbf{LoF}      & \textbf{LP} & \textbf{LoF(LP)} & \textbf{MSE(L)} \\
1 & 1    & 0    & 0    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{3} & 3 & 2 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{100.00} & 91.95  & 18.79  \\
2 & 0    & 1    & 0    & \multicolumn{1}{|r}{3} & 1 & 2 & \multicolumn{1}{|r}{4} & 1 & 3 & \multicolumn{1}{|r}{13} & 0  & \multicolumn{1}{|r}{0.00}   & 100.00 & 0.00   \\
3 & 0    & 0    & 1    & \multicolumn{1}{|r}{0} & 3 & 3 & \multicolumn{1}{|r}{0} & 0 & 8 & \multicolumn{1}{|r}{1}  & 12 & \multicolumn{1}{|r}{0.00}   & 0.00   & 100.00 \\
4 & 1/3 & 1/3 & 1/3    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{0} & 5 & 3 & \multicolumn{1}{|r}{9}  & 4  & \multicolumn{1}{|r}{76.38}  & 86.50  & 29.72  \\
5 & 0.5  & 0.5  & 0    & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{2} & 4 & 2 & \multicolumn{1}{|r}{11} & 2  & \multicolumn{1}{|r}{91.18}  & 95.72  & 17.55  \\
6 & 0.5  & 0    & 0.5  & \multicolumn{1}{|r}{2} & 1 & 3 & \multicolumn{1}{|r}{0} & 5 & 3 & \multicolumn{1}{|r}{8}  & 5  & \multicolumn{1}{|r}{72.48}  & 79.24  & 34.81  \\
7 & 0    & 0.5  & 0.5  & \multicolumn{1}{|r}{0} & 3 & 3 & \multicolumn{1}{|r}{1} & 4 & 3 & \multicolumn{1}{|r}{8}  & 5  & \multicolumn{1}{|r}{37.45}  & 78.53  & 49.34  \\
8 & 0.5  & 0.25 & 0.25 & \multicolumn{1}{|r}{3} & 1 & 2 & \multicolumn{1}{|r}{3} & 2 & 3 & \multicolumn{1}{|r}{10} & 3  & \multicolumn{1}{|r}{93.63}  & 91.72  & 23.68 
\end{tabular}
}
\end{table}